---
title: "DataWrangling"
author: "Matt"
date: "1/31/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Skills checklist

- can input files in .txt, .csv, and .xlsx into data.frames or data.tables
- can rename columns
- can convert columns into desired class (e.g., a factor vs not a factor)
- 

### wide to long

Rosenbaum, Mama, Algom, (2017) investigated whether Stroop effects change depending on whether subjects are sitting or standing. A tutorial on this dataset is [https://crumplab.github.io/statisticsLab/lab-10-factorial-anova.html#does-standing-up-make-you-focus-more](https://crumplab.github.io/statisticsLab/lab-10-factorial-anova.html#does-standing-up-make-you-focus-more).

The data file contains 4 columns. Each row is for one subject. Each row contains 4 means, for the incongruent and congruent stroop conditions, both when subjects were sitting and standing

**pre-processing**

 - load the data
 - convert from wide format to long format (each row should only contain one dependent measure per subject)

```{r}
library(data.table)
all_data <- fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/stroop_stand.csv")
```

**Analysis questions**

  - use dplyr to calculate the means for congruent vs. incongruent, collapsing over the sitting/standing variable
  - use dplyr to calculate the standard error for congruent vs. incongruent, collapsing over the sitting/standing variable
  - use dplyr to calculate the means for sitting/standing variable collapsing over congruency
  - use dplyr to calculate the standard error for sitting/standing variable collasping over congruency
  - use dplyr to calculate the condition means for each cell in the 2x2 design.
  - use dplyr to calculate the condition standard errors for each cell in the 2x2 design.
  - According to the means, was the average Stroop effect different depending on whether subjects were sitting or standing?

-----

### Means at multiple levels

```{r}
all_data <- fread("https://github.com/CrumpLab/statisticsLab/raw/master/data/exp1_BehmerCrumpAPP.csv")
```

**Analysis Questions**

a. Which subject had the worst overall accuracy?
b. How many subjects had an overall accuracy worse than 80%?
c. How many subjects had an accuracy of less than 80% in any one of the six conditions?
d. Create a scatterplot showing each subjects mean overall accuracy and mean overall IKSI (interkeystroke interval)



-----

### Data-wrangling text

Obtain a reasonably sized piece of text (at least 250 words or more) and load into R as a string variable

**Pre-processing**

 - convert all characters to lowercase, remove all punctuation, special characters,and numbers so the only characters left are the letters from a-z
 
**Analysis questions**

 - create a table with a letter frequency distribution (a count of how many times each letter appears in the text)
 - Letters appear in different positions of words. Create a table with letter frequency distrbibutions as a function of letter position in a word. For example, you will have a distribution for all first position letters, all 2nd position letters, and so on. Don't worry about going past 9 positions.


