---
title: "Statistics"
author: "Matt"
date: "1/31/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

## t-tests

### one sample t-test

Set `mu` to test against a population value

```{r}
x <- c(1,4,3,2,3,4,3,2,3,2,3,4)

#non-directional
t.test(x, mu = 2)
t.test(x, mu = 2, alternative = "two.sided") # the default

# directional
t.test(x, mu = 2, alternative = "greater")
t.test(x, mu = 2, alternative = "less")
```

### paired-sample t-test

Set `paired=TRUE`

```{r}
x <- c(1,4,3,2,3,4,3,2,3,2,3,4)
y <- c(3,2,5,4,3,2,3,2,1,2,2,3)

#non-directional
t.test(x, y, paired=TRUE)
t.test(x, y, paired=TRUE, alternative = "two.sided") # the default

# directional
t.test(x, y, paired=TRUE, alternative = "greater")
t.test(x, y, paired=TRUE, alternative = "less")
```

### Independent sample t-test

Note: Default is Welch test, set `var.equal=TRUE` for t-test

```{r}
x <- c(1,4,3,2,3,4,3,2,3,2,3,4)
y <- c(3,2,5,4,3,2,3,2,1,2,2,3)

#non-directional
t.test(x, y, var.equal=TRUE)
t.test(x, y, var.equal=TRUE, alternative = "two.sided") # the default

# directional
t.test(x, y, var.equal=TRUE, alternative = "greater")
t.test(x, y, var.equal=TRUE, alternative = "less")
```

### printing 

```{r}
library(broom)

t_results <- t.test(x, y, var.equal=TRUE)
knitr::kable(tidy(t_results))

```

### writing

The contents of R variables can be written into R Markdown documents. 

t(`r round(t_results$parameter, digits=2)`) =`r round(t_results$statistic, digits=2)`, p = `r round(t_results$p.value, digits=3)`

```{r}
# write your own function
report_t <- function(x){
  return(paste(c("t(",round(x$parameter, digits=2),") ",
        "= ",round(x$statistic, digits=2),", ",
        "p = ",round(x$p.value, digits=3)), collapse=""))
}
report_t(t_results)
t_results
```


The results of my t-test were `r report_t(t_results)`.


## ANOVA

Requires data to be in long-format.

```{r}
# example creation of 2x2 data 
Subjects <- rep(1:10,each=4)
Factor1 <- rep(rep(c("A","B"), each = 2), 10)
Factor2 <- rep(rep(c("1","2"), 2), 10)
DV <- rnorm(40,0,1)

all_data <- data.frame(Subjects = as.factor(Subjects),
                       DV,
                       Factor1,
                       Factor2)
```

### between-subjects 1 factor

```{r}

# run anova
aov_out <- aov(DV~Factor1, all_data)

# summary
summary(aov_out)

# kable xtable printing
library(xtable)
knitr::kable(xtable(summary(aov_out)))

# print means
print(model.tables(aov_out,"means"), format="markdown")
```

### between-subjects 2 factor

```{r}

# run anova
aov_out <- aov(DV~Factor1*Factor2, all_data)

# summary
summary(aov_out)

# kable xtable printing
library(xtable)
knitr::kable(xtable(summary(aov_out)))

# print means
print(model.tables(aov_out,"means"), format="markdown")
```

### repeated-measures 1 factor

```{r}

# run anova
aov_out <- aov(DV~Factor1 + Error(Subjects/Factor1), all_data)

# summary
summary(aov_out)

# kable xtable printing
library(xtable)
knitr::kable(xtable(summary(aov_out)))

# print means
print(model.tables(aov_out,"means"), format="markdown")
```


### repeated-measures 2 factor

```{r}

# run anova
aov_out <- aov(DV~Factor1*Factor2 + Error(Subjects/(Factor1*Factor2)), all_data)

# summary
summary(aov_out)

# kable xtable printing
library(xtable)
knitr::kable(xtable(summary(aov_out)))

# print means
print(model.tables(aov_out,"means"), format="markdown")
```

## Linear Regression

```{r}
lm(DV~Factor1, all_data)
summary(lm(DV~Factor1, all_data))
```

```{r}
lm(DV~Factor1+Factor2, all_data)
summary(lm(DV~Factor1+Factor2, all_data))
```

```{r}
lm(DV~Factor1*Factor2, all_data)
summary(lm(DV~Factor1*Factor2, all_data))
```



### papaja

```{r}
library(papaja)

apa_stuff <- apa_print.aov(aov_out)
```

The main effect for factor 1 was, `r apa_stuff$statistic$Factor1`. The main effect for factor 2 was, `r apa_stuff$statistic$Factor2`. The interaction was, `r apa_stuff$statistic$Factor1_Factor2`

## Randomization Test

```{r}
A <- c(1,2,3,4,5,6,7,8,9,10)
B <- c(2,4,6,8,10,12,14,16,18,20)

all <- c(A,B)

mean_difference <- c()
for(i in 1:10000){
  shuffle <- sample(all)
  newA <- shuffle[1:10]
  newB <- shuffle[11:20]
  mean_difference[i] <- mean(newB)-mean(newA)
}

observed <- mean(B)-mean(A)
length(mean_difference[mean_difference >= observed])/10000

library(EnvStats)

twoSamplePermutationTestLocation(x=B, y=A, fcn = "mean", 
                                 alternative = "greater", 
                                 mu1.minus.mu2 = 0, 
                                 paired = FALSE, 
                                 exact = FALSE, 
                                 n.permutations = 10000, 
                                 seed = NULL, 
                                 tol = sqrt(.Machine$double.eps))

```

## Correlation

```{r}
x <- rnorm(10,0,1)
y <- rnorm(10,0,1)
cor(x,y)

ranks1 <- sample(1:10) 
ranks2 <- sample(1:10)
cor(ranks1,ranks2, method = "spearman")
```

## Chi-square

```{r}
M <- as.table(rbind(c(762, 327, 468), c(484, 239, 477)))
dimnames(M) <- list(gender = c("F", "M"),
                    party = c("Democrat","Independent", "Republican"))
(Xsq <- chisq.test(M))
```

## binomial test

```{r}
number_of_successes <- 100
trials <- 150
binom.test(x=number_of_successes, n=trials, p = 0.5,
           alternative = "two.sided",
           conf.level = 0.95)
```

